RDF Validation Requirements - Evaluation and Logical Underpinning

### 2014-12-26 paper/RDF validation requirements/RDF validation requirements.tex
added mapping tables

\newenvironment{gcotable}{
  \scriptsize
  \sffamily
  \vspace{0.3cm}
	\begin{center}
  \begin{tabular}{l|l|l|l|l|l|l}
  \hline
  \textbf{c. type} & \textbf{context class} & \textbf{left r. list} & \textbf{right r. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
  \hline

@@

%\tb{ToDo: check DL}

@@

%\tb{ToDo: check DL}

@@

\begin{itemize}
	\item \ms{skos:Concept} resources can only be members of the listed controlled vocabularies
\end{itemize}
\er{This sort of DL-syntax (with semicolon) must be rechecked. Note that in DL, it does not have to correspond directly to names of things in machine-readable syntax. It is fine to say BookSubjects rather than :BookSubjects for the sake of simplicity }
\begin{DL}
$A \equiv skos:Concept \sqcap B$ \\
$B \equiv \forall skos:inScheme . C$ \\
$C \equiv skos:ConceptScheme \sqcap ( \{:BookSubjects\} \sqcup \{:BookTopics\} \sqcup \{:BookCategories\} )$ \\
\end{DL}

@@

@@

\begin{DL}
NOT EXPRESSIBLE IN DL!!!
\end{DL}

@@

@@

@@

\begin{DL}
1-23 $\equiv$ \{'1'\} $\sqcup$ \{'2'\} $\sqcup$ ... $\sqcup$ \{'23'\} \\
$\top$ $\sqsubseteq$ $\forall$ position . 1-23 \\
This is just an auxiliary construction and is not sufficient for general cases, so: NOT EXPRESSIBLE IN DL!!!
\end{DL}

\begin{gcotable}
class & AtLeast1 & - & - & xsd:nonNegativeInteger & $\geq$ & 1 \\
class & AtMost23 & - & - & xsd:nonNegativeInteger & $\leq$ & 23 \\
class & NumberPlayers... & - & - & AtLeast1, AtMost23 & $\sqcap$ & - \\
property & $\top$ & position & - & NumberPlayers... & range & - \\
\end{gcotable}

or:

@@

or:

\begin{gcotable}
class & 1-23 & - & - & \{'1'\}, \{'2'\}, ..., \{'23'\} & $\sqcup$ & - \\
property & $\top$ & position & - & 1-23 & range& - \\
\end{gcotable}

@@

\begin{DL}
I'm not sure about the meaning of this constraint, but for me this seams like e.g. the range of a role... but if this is the case, take into account the consequences of reasoning with OWA...
\end{DL}

@@

\begin{DL}
I'm not sure about the meaning of this constraint, but for me this seams like e.g. the domain of a role... but if this is the case, take into account the consequences of reasoning with OWA...
\end{DL}

@@

\begin{DL}
NOT EXPRESSIBLE IN DL!!!
\end{DL}

@@

\begin{DL}
NOT EXPRESSIBLE IN DL!!!
\end{DL}

@@

\begin{DL}
NOT EXPRESSIBLE IN DL!!!
\end{DL}

@@

\begin{DL}
NOT EXPRESSIBLE IN DL!!!
\end{DL}

@@

\begin{DL}
NOT EXPRESSIBLE IN DL!!!
\end{DL}

@@

\begin{itemize}
	\item check if there are no html tags included in literals (of specific data properties within the context of specific classes)
\end{itemize}

@@

\begin{DL}
NOT EXPRESSIBLE IN DL!!!
\end{DL}

@@

\begin{DL}
$\exists hasAncestor.Person \sqsubseteq Person$ \\
(same as the definition of domains)
\end{DL}

@@

\begin{DL}
Not sure about the meaning of this constraint...
\end{DL}

@@

\begin{DL}
- condition: universal quantification on object and data properties\\
\ \ \ \ same as above!?\\
- condition: existential quantification on object and data properties\\
\ \ \ \ same as above!?\\
- condition: if specific properties are present, then specific other properties also have to be present
\ \ \ \ same as super properties, so if a property (role or attribute) is used, all super properties are also present
\end{DL}

@@

\begin{DL}
Not sure about the meaning of this constraint...
\end{DL}

>>>>

\newenvironment{gcotable}{
  \scriptsize
  \sffamily
  \vspace{0.3cm}
	\begin{center}
  \begin{tabular}{l|l|l|l|l|l|l}
  \hline
  \textbf{c. type} & \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
  \hline

@@

\begin{gcotable}
property & Jedi & laserSwordColor & - & LaserSwordColor & $\exists$ & - \\
class & LaserSwordColor & - & - & $\{blue\}$, $\{green\}$, $\{white\}$ & $\sqcup$ & - \\
\end{gcotable}

@@

\begin{gcotable}
property & $\exists$ laserSwordColor . LaserSwordColor & laserSwordColor & - & LaserSwordColor & $\exists$ & - \\
class & Jedi & - & - & $\exists$ laserSwordColor . LaserSwordColor & $\neg$ & - \\
class & LaserSwordColor & - & - & $\{blue\}$, $\{green\}$, $\{white\}$ & $\sqcup$ & - \\
\end{gcotable}

@@

\begin{itemize}
	\item \ms{skos:Concept} resources can only be members of the listed controlled vocabularies
\end{itemize}
%\er{This sort of DL-syntax (with semicolon) must be rechecked. Note that in DL, it does not have to correspond directly to names of things in machine-readable syntax. It is fine to say BookSubjects rather than :BookSubjects for the sake of simplicity }
\begin{DL}
$A \equiv Concept \sqcap B$ \\
$B \equiv \forall inScheme . C$ \\
$C \equiv ConceptScheme \sqcap ( \{BookSubjects\} \sqcup \{BookTopics\} \sqcup \{BookCategories\} )$
\end{DL}

\begin{gcotable}
property & $\top$ & subject & - & A & range & - \\
class & A & - & - & Concept, B & $\sqcap$ & - \\
property & B & inScheme & - & C & $\forall$ & - \\
class & C & - & - & ConceptScheme, D & $\sqcap$ & - \\
class & D & - & - & $\{BookSubjects\}$, $\{BookTopics\}$, $\{BookCategories\}$ & $\sqcup$ & - \\
\end{gcotable}

@@

\begin{ex}
# OWL 2 DL (functional-style syntax)
# ----------------------------------
Declaration( Datatype( :SSN ) ) 
DatatypeDefinition( 
    :SSN
    DatatypeRestriction( xsd:string xsd:pattern "[0-9]{3}-[0-9]{2}-[0-9]{4}" ) )     
DataPropertyRange( :hasSSN :SSN ) 
\end{ex}

\begin{ex}
# OWL 2 DL (turtle syntax)
# ------------------------
:SSN 
    a rdfs:Datatype ;
    owl:equivalentClass [
        a rdfs:Datatype ;
        owl:onDatatype xsd:string ;
        owl:withRestrictions ( 
            [ xsd:pattern "[0-9]{3}-[0-9]{2}-[0-9]{4}" ] ) ] .
:hasSSN rdfs:range :SSN .
\end{ex}

@@

\begin{DL}
Not expressible in DL!
\end{DL}

@@

There are multiple use cases associated with the requirement to match literals according to given patterns.

Luke's droids can only have the numbers "R2-D2" or "C-3PO".
The universal restriction part of this constraint can be expressed by OWL 2 DL:
\ms{LukesDroids $\sqsubseteq$ $\forall$ droidNumber . DroidNumber}.
The restriction of the datatype \ms{DroidNumber}, however, cannot be expressed in DL, but OWL 2 DL can be used anyway:

\begin{ex}
DroidNumber 
    a rdfs:Datatype ;
    owl:equivalentClass [
        a rdfs:Datatype ;
        owl:onDatatype xsd:string ;
        owl:withRestrictions ( 
            [ xsd:pattern "R2-D2|C-3PO" ] ) ] .
\end{ex}

The second axiom defines \ms{DroidNumber} as an abbreviation for a datatype restriction on \ms{xsd:string}. 
The first axiom explicitly declares \ms{DroidNumber} to be a datatype. 
The datatype \ms{DroidNumber} can be used just like any other datatype like in the universal restriction above.
The literal pattern matching constraint validates \ms{DroidNumber} literals according to the stated regular expression causing a constraint violation for the triples 
\ms{Luke hasDroid Droideka} and \ms{Droideka droidNumber "Droideka"\textasciicircum{}\textasciicircum{}DroidNumber}, 
but not for the triples \ms{Luke hasDroid R2-D2} and \ms{R2-D2 droidNumber "R2-D2"\textasciicircum{}\textasciicircum{}DroidNumber}.

\begin{DL}
Not expressible in DL!
\end{DL}

\begin{gcotable}
property & LukesDroids & droidNumber & - & DroidNumber & $\forall$ & - \\
class & DroidNumber & - & - & xsd:string & xsd:pattern & 'R2-D2$|$C-3PO' \\
\end{gcotable}

@@

\begin{DL}
Not expressible in DL!
\end{DL}

\begin{gcotable}
property & $\top$ & dc:subject & dc:subject & xsd:string & $\neq$ & - \\
\end{gcotable}

@@

%\begin{DL}
%1-23 $\equiv$ \{'1'\} $\sqcup$ \{'2'\} $\sqcup$ ... $\sqcup$ \{'23'\} \\
%$\top$ $\sqsubseteq$ $\forall$ position . 1-23 \\
%\end{DL}

\begin{DL}
Not expressible in DL!
\end{DL}

%\begin{gcotable}
%class & AtLeast1 & - & - & xsd:nonNegativeInteger & $\geq$ & 1 \\
%class & AtMost23 & - & - & xsd:nonNegativeInteger & $\leq$ & 23 \\
%class & NumberPlayers... & - & - & AtLeast1, AtMost23 & $\sqcap$ & - \\
%property & $\top$ & position & - & NumberPlayers... & range & - \\
%\end{gcotable}
%
%or:

@@

%or:
%
%\begin{gcotable}
%class & 1-23 & - & - & \{'1'\}, \{'2'\}, ..., \{'23'\} & $\sqcup$ & - \\
%property & $\top$ & position & - & 1-23 & range& - \\
%\end{gcotable}

@@

I'm not sure about the meaning of this constraint, but for me this seams like e.g. the range of a role... but if this is the case, take into account the consequences of reasoning with OWA...

@@

I'm not sure about the meaning of this constraint, but for me this seams like e.g. the domain of a role... but if this is the case, take into account the consequences of reasoning with OWA...

@@

\begin{DL}
Not expresible in DL!
\end{DL}

\begin{gcotable}
class & Jedi & laserSwordColor & - & - & default values & 'blue' \\
class & Jedi & numberLaserSwords & - & - & default values & '1' \\
class & Sith & laserSwordColor & - & - & default values & 'red' \\
class & Sith & numberLaserSwords & - & - & default values & '2' \\
\end{gcotable}

@@

\begin{DL}
Not Expressible in DL!
\end{DL}

@@

Only English names are allowed.

\begin{DL}
Not expressible in DL!
\end{DL}

\begin{gcotable}
class & $\top$ & name & - & - & language tag & 'en' \\
\end{gcotable}

@@

\begin{DL}
Not expressible in DL!
\end{DL}

\begin{gcotable}
class & $\top$ & foaf:name & - & - & language tag exact cardinality & 1 \\
\end{gcotable}

@@

check if literals of the property \ms{p} do not include whitespaces

\begin{DL}
Not expressible in DL!
\end{DL}

\begin{gcotable}
class & $\top$ & p & - & - & whitespace & - \\
\end{gcotable}

@@

\begin{itemize}
	\item check if there are no HTML tags included in literals (of specific data properties within the context of specific classes)
\end{itemize}

@@

check if literals of the property \ms{p} of the class \ms{C} do not include HTML tags

\begin{DL}
Not expressible in DL!
\end{DL}

\begin{gcotable}
class & c & p & - & - & HTML & - \\
\end{gcotable}

@@

\begin{DL}
$\exists hasAncestor.Person \sqsubseteq Person$ \\
\end{DL}

Same as the definition of domains.

@@

@@

@@

Property \ms{p} is recommended to use within the context of the class \ms{C}.

\begin{DL}
Not expressible in DL!
\end{DL}

\begin{gcotable}
class & C & p & - & - & recommended & - \\
\end{gcotable}
